---
title: "Manuscript Figures: Apple Orchard Eddy-Covariance Study"
author: "Kevin Postma"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.align = "center"
)
```

# Purpose

This script generates publication-quality versions of:

- **Figure 2:** cumulative growing-season precipitation in 2024 and 2025;
- **Figure 3:** chained cumulative NECB with weekly paired NEE differences.

Figure 2 is derived from `Fluxes_Drivers3_2024_2025.xlsx`. Figure 3 is derived from the final annual U50 daily processing CSVs so that it reflects the winter positive/negative reviews and the seeded REddyProc processing runs.

```{r packages_paths_helpers}
# ============================================================
# 0) PACKAGES
# ============================================================

required_packages <- c(
  "dplyr", "tidyr", "readr", "readxl", "lubridate",
  "ggplot2", "patchwork", "purrr", "stringr", "tibble", "here"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Install the following packages before knitting this script: ",
    paste(missing_packages, collapse = ", ")
  )
}

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(readr)
  library(readxl)
  library(lubridate)
  library(ggplot2)
  library(patchwork)
  library(purrr)
  library(stringr)
  library(tibble)
  library(here)
})

# ============================================================
# 1) PROJECT PATHS
# ============================================================

project_root <- here::here()
processed_root <- file.path(project_root, "data", "processed")

figure_main_dir <- file.path(project_root, "figures", "manuscript")
figure_support_dir <- file.path(
  processed_root, "publication", "figure_support"
)

walk(
  c(figure_main_dir, figure_support_dir),
  ~ dir.create(.x, recursive = TRUE, showWarnings = FALSE)
)

figure2_workbook <- file.path(
  processed_root,
  "Fluxes_Drivers3_2024_2025.xlsx"
)

if (!file.exists(figure2_workbook)) {
  stop(
    "Could not find Fluxes_Drivers3_2024_2025.xlsx at: ",
    figure2_workbook,
    "\nOpen the apple-orchard-eddy-covariance RStudio project and try again."
  )
}

message("Project root: ", project_root)
message("Figure output directory: ", figure_main_dir)

# ============================================================
# 2) STUDY METADATA
# ============================================================

years_use <- c(2024L, 2025L)

# Phenology-based growing-season dates.
gs_dates <- tribble(
  ~year, ~gs_start,             ~gs_end,
  2024L, as.Date("2024-03-28"), as.Date("2024-10-31"),
  2025L, as.Date("2025-04-15"), as.Date("2025-10-31")
)

# Harvested-fruit carbon export and plotting dates.
# Change HarvestDate here if a different accounting date is preferred.
harvest_C <- tribble(
  ~year, ~Orchard, ~Harvest_gC_m2, ~HarvestDate,
  2024L, "LD", 177, as.Date("2024-11-01"),
  2024L, "MD", 337, as.Date("2024-11-01"),
  2025L, "LD",  68, as.Date("2025-11-01"),
  2025L, "MD", 331, as.Date("2025-11-01")
)

# Publication colors.
# LD is blue and MD is orange. Figure 2 uses blue for 2024 and orange for 2025.
orchard_colors <- c(
  "LD" = "#0072B2",
  "MD" = "#D55E00"
)

year_colors <- c(
  "2024" = "#0072B2",
  "2025" = "#E69F00"
)

# ============================================================
# 3) GENERAL HELPERS
# ============================================================

safe_sum <- function(x) {
  if (all(is.na(x))) NA_real_ else sum(x, na.rm = TRUE)
}

parse_date_safely <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))
  if (is.numeric(x)) return(as.Date(x, origin = "1899-12-30"))

  x_chr <- trimws(as.character(x))
  out <- suppressWarnings(lubridate::ymd(x_chr, quiet = TRUE))

  unresolved <- is.na(out) & nzchar(x_chr)
  if (any(unresolved)) {
    out[unresolved] <- suppressWarnings(
      lubridate::mdy(x_chr[unresolved], quiet = TRUE)
    )
  }

  as.Date(out)
}

parse_logical_flag <- function(x, label = "logical column") {
  if (is.logical(x)) return(x)

  x_chr <- tolower(trimws(as.character(x)))

  out <- case_when(
    is.na(x_chr) | x_chr == "" ~ NA,
    x_chr %in% c("true", "t", "1", "yes", "y") ~ TRUE,
    x_chr %in% c("false", "f", "0", "no", "n") ~ FALSE,
    TRUE ~ NA
  )

  invalid <- !is.na(x_chr) & x_chr != "" & is.na(out)

  if (any(invalid)) {
    stop(
      label,
      " contains values that cannot be interpreted as TRUE or FALSE: ",
      paste(sort(unique(x_chr[invalid])), collapse = ", ")
    )
  }

  out
}

find_one_file <- function(root, pattern, label) {
  hits <- list.files(
    root,
    pattern = pattern,
    full.names = TRUE,
    recursive = TRUE,
    ignore.case = TRUE
  )

  hits <- hits[!grepl("(^|/|\\\\)~\\$", hits)]

  if (length(hits) == 0) {
    stop("Could not find ", label, " under: ", root)
  }

  if (length(hits) > 1) {
    stop(
      "Found more than one possible ", label, ":\n",
      paste(hits, collapse = "\n")
    )
  }

  hits[[1]]
}

resolve_daily_csv <- function(year_use) {
  year_root <- file.path(processed_root, as.character(year_use))

  find_one_file(
    year_root,
    paste0(
      "^Daily_LD_MD_Fluxes_plus_Drivers_plus_VPD_",
      year_use,
      "_U50\\.csv$"
    ),
    paste0("final daily U50 flux-and-driver file for ", year_use)
  )
}

resolve_results_table <- function(year_use) {
  year_root <- file.path(processed_root, as.character(year_use))

  find_one_file(
    year_root,
    paste0("^Results_table_", year_use, "_U50\\.csv$"),
    paste0("final Results_table for ", year_use)
  )
}

publication_theme <- function(base_size = 11) {
  theme_bw(base_size = base_size, base_family = "Helvetica") +
    theme(
      text = element_text(family = "Helvetica", color = "black"),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10, color = "black"),
      legend.title = element_text(size = 11, face = "bold"),
      legend.text = element_text(size = 10),
      strip.text = element_text(size = 11, face = "bold"),
      plot.tag = element_text(size = 12, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.margin = margin(6, 8, 6, 8)
    )
}

save_publication_figure <- function(
  plot,
  basename,
  width,
  height
) {
  png_path <- file.path(figure_main_dir, paste0(basename, ".png"))
  tiff_path <- file.path(figure_main_dir, paste0(basename, ".tiff"))
  pdf_path <- file.path(figure_main_dir, paste0(basename, ".pdf"))

  ggsave(
    filename = png_path,
    plot = plot,
    width = width,
    height = height,
    units = "in",
    dpi = 300,
    bg = "white"
  )

  ggsave(
    filename = tiff_path,
    plot = plot,
    width = width,
    height = height,
    units = "in",
    dpi = 600,
    compression = "lzw",
    bg = "white"
  )

  pdf_device <- if (capabilities("cairo")) {
    grDevices::cairo_pdf
  } else {
    grDevices::pdf
  }

  ggsave(
    filename = pdf_path,
    plot = plot,
    width = width,
    height = height,
    units = "in",
    device = pdf_device,
    bg = "white"
  )

  created <- c(png_path, tiff_path, pdf_path)
  missing <- created[!file.exists(created)]

  if (length(missing) > 0) {
    stop("One or more figure files were not created: ", paste(missing, collapse = ", "))
  }

  message("Saved figure files: ", paste(created, collapse = " | "))
}
```

# Load final daily flux data for Figure 3

```{r load_final_daily_fluxes}
# ============================================================
# 4) FINAL DAILY U50 FLUX FILES FOR FIGURE 3
# ============================================================

daily_paths <- set_names(
  map_chr(years_use, resolve_daily_csv),
  as.character(years_use)
)

walk2(
  names(daily_paths),
  unname(daily_paths),
  ~ message("Final daily processing file for ", .x, ": ", .y)
)

read_final_daily_year <- function(path, year_use) {
  x <- read_csv(
    path,
    show_col_types = FALSE,
    na = c("", "NA", "NaN")
  )

  required <- c("Date", "LD_NEE_gC_day", "MD_NEE_gC_day")
  missing <- setdiff(required, names(x))

  if (length(missing) > 0) {
    stop(
      "The final daily file for ", year_use,
      " is missing: ", paste(missing, collapse = ", "),
      "\nFile: ", path
    )
  }

  x <- x %>%
    mutate(
      Date = parse_date_safely(Date),
      LD_NEE_gC_day = parse_number(as.character(LD_NEE_gC_day)),
      MD_NEE_gC_day = parse_number(as.character(MD_NEE_gC_day))
    ) %>%
    filter(!is.na(Date))

  outside_year <- x %>%
    filter(lubridate::year(Date) != year_use)

  if (nrow(outside_year) > 0) {
    message(
      "Removing ", nrow(outside_year),
      " boundary-date row(s) from the ", year_use,
      " daily file: ",
      paste(sort(unique(as.character(outside_year$Date))), collapse = ", ")
    )
  }

  x %>%
    filter(lubridate::year(Date) == year_use) %>%
    transmute(
      year = as.integer(year_use),
      Date,
      LD_NEE_gC_day,
      MD_NEE_gC_day,
      source_file = basename(path)
    )
}

final_daily <- map2_dfr(
  unname(daily_paths),
  years_use,
  read_final_daily_year
) %>%
  arrange(Date)

if (anyDuplicated(final_daily$Date) > 0) {
  duplicate_dates <- final_daily %>%
    count(Date) %>%
    filter(n > 1)

  print(duplicate_dates, n = Inf)
  stop("Duplicate dates remain in the final daily files.")
}

write_csv(
  tibble(
    year = as.integer(names(daily_paths)),
    daily_source_path = unname(daily_paths)
  ),
  file.path(figure_support_dir, "Figure_3_daily_source_manifest.csv")
)

# Compare daily-file annual sums with the authoritative half-hour summaries.
# Small differences can occur because the daily products apply a daily coverage rule.
results_paths <- set_names(
  map_chr(years_use, resolve_results_table),
  as.character(years_use)
)

results_table_all <- map_dfr(
  years_use,
  function(year_use) {
    read_csv(results_paths[[as.character(year_use)]], show_col_types = FALSE) %>%
      mutate(
        Year = as.integer(Year),
        Orchard = toupper(as.character(Orchard)),
        Period = as.character(Period)
      )
  }
)

figure3_validation <- final_daily %>%
  group_by(year) %>%
  summarise(
    LD = safe_sum(LD_NEE_gC_day),
    MD = safe_sum(MD_NEE_gC_day),
    .groups = "drop"
  ) %>%
  pivot_longer(c(LD, MD), names_to = "Orchard", values_to = "Daily_NEE") %>%
  left_join(
    results_table_all %>%
      filter(Period == "Full-year") %>%
      select(year = Year, Orchard, Official_NEE = NEE),
    by = c("year", "Orchard")
  ) %>%
  mutate(
    Difference = Daily_NEE - Official_NEE,
    Absolute_difference = abs(Difference)
  )

print(figure3_validation, n = Inf)

write_csv(
  figure3_validation,
  file.path(
    figure_support_dir,
    "Figure_3_daily_vs_official_annual_NEE_validation.csv"
  )
)

if (any(figure3_validation$Absolute_difference > 10, na.rm = TRUE)) {
  stop(
    "A final daily CSV differs from the official annual NEE by more than 10 g C m-2. ",
    "Review Figure_3_daily_vs_official_annual_NEE_validation.csv."
  )
}

if (any(figure3_validation$Absolute_difference > 1.1, na.rm = TRUE)) {
  warning(
    "Figure 3 daily trajectories differ slightly from one or more official annual totals. ",
    "This is expected when low-coverage daily values are omitted; see the validation CSV."
  )
}
```

# Figure 2: cumulative precipitation

```{r figure_2}
# ============================================================
# 5) FIGURE 2: CUMULATIVE GROWING-SEASON PRECIPITATION
# ============================================================

figure2_raw <- read_excel(
  figure2_workbook,
  sheet = 1,
  .name_repair = "unique"
)

required_figure2_columns <- c(
  "Date", "year", "P_RAIN_tower_mm"
)

missing_figure2_columns <- setdiff(
  required_figure2_columns,
  names(figure2_raw)
)

if (length(missing_figure2_columns) > 0) {
  stop(
    "Fluxes_Drivers3_2024_2025.xlsx is missing columns required for Figure 2: ",
    paste(missing_figure2_columns, collapse = ", ")
  )
}

figure2_daily <- figure2_raw %>%
  mutate(
    Date = parse_date_safely(Date),
    year = as.integer(parse_number(as.character(year))),
    P_RAIN_tower_mm = parse_number(as.character(P_RAIN_tower_mm))
  ) %>%
  filter(
    year %in% years_use,
    !is.na(Date)
  ) %>%
  left_join(gs_dates, by = "year") %>%
  mutate(
    GrowingSeason_calc = Date >= gs_start & Date <= gs_end
  ) %>%
  filter(GrowingSeason_calc) %>%
  arrange(year, Date)

# Validate the workbook's GrowingSeason field when it is present.
if ("GrowingSeason" %in% names(figure2_raw)) {
  gs_check <- figure2_raw %>%
    mutate(
      Date = parse_date_safely(Date),
      year = as.integer(parse_number(as.character(year))),
      GrowingSeason_source = parse_logical_flag(
        GrowingSeason,
        label = "GrowingSeason in Fluxes_Drivers3_2024_2025.xlsx"
      )
    ) %>%
    filter(year %in% years_use, !is.na(Date)) %>%
    left_join(gs_dates, by = "year") %>%
    mutate(
      GrowingSeason_calc = Date >= gs_start & Date <= gs_end
    ) %>%
    filter(
      !is.na(GrowingSeason_source),
      GrowingSeason_source != GrowingSeason_calc
    )

  if (nrow(gs_check) > 0) {
    print(gs_check, n = Inf)
    stop(
      "GrowingSeason values in Fluxes_Drivers3_2024_2025.xlsx do not match ",
      "the documented phenology dates."
    )
  }
}

missing_rain <- figure2_daily %>%
  filter(!is.finite(P_RAIN_tower_mm))

if (nrow(missing_rain) > 0) {
  print(missing_rain %>% select(year, Date, P_RAIN_tower_mm), n = Inf)
  stop(
    "Figure 2 contains missing tower-precipitation values during the growing season. ",
    "Do not silently replace these with zero."
  )
}

figure2_support <- figure2_daily %>%
  group_by(year) %>%
  mutate(
    Cumulative_precipitation_mm = cumsum(P_RAIN_tower_mm)
  ) %>%
  ungroup() %>%
  transmute(
    Year = factor(year, levels = years_use),
    Date,
    PlotDate = as.Date(paste0("2000-", format(Date, "%m-%d"))),
    Daily_precipitation_mm = P_RAIN_tower_mm,
    Cumulative_precipitation_mm
  )

figure2_totals <- figure2_support %>%
  group_by(Year) %>%
  summarise(
    Growing_season_days = n_distinct(Date),
    Total_precipitation_mm = sum(Daily_precipitation_mm),
    .groups = "drop"
  )

print(figure2_totals, n = Inf)

write_csv(
  figure2_support,
  file.path(figure_support_dir, "Figure_2_cumulative_precipitation_support.csv")
)

write_csv(
  figure2_totals,
  file.path(figure_support_dir, "Figure_2_precipitation_totals.csv")
)

p_figure2 <- ggplot(
  figure2_support,
  aes(
    x = PlotDate,
    y = Cumulative_precipitation_mm,
    color = Year,
    group = Year
  )
) +
  geom_step(linewidth = 1.15, direction = "hv") +
  scale_color_manual(
    values = year_colors,
    breaks = c("2024", "2025")
  ) +
  scale_x_date(
    limits = c(as.Date("2000-03-20"), as.Date("2000-11-10")),
    date_breaks = "1 month",
    date_labels = "%b",
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.06))
  ) +
  labs(
    x = "Month",
    y = "Cumulative precipitation (mm)",
    color = "Year"
  ) +
  publication_theme(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    legend.position = c(0.78, 0.30),
    legend.background = element_rect(
      fill = "white",
      color = NA
    )
  )

print(p_figure2)

save_publication_figure(
  p_figure2,
  basename = "Figure_2_Cumulative_Growing_Season_Precipitation_2024_2025",
  width = 7.2,
  height = 4.8
)
```

# Figure 3: chained cumulative NECB and weekly NEE difference

```{r figure_3}
# ============================================================
# 6) FIGURE 3: CHAINED CUMULATIVE NECB + WEEKLY NEE DIFFERENCE
# ============================================================

# Orchard-specific daily NEE trajectories. Each orchard retains all of its
# available daily values; the upper panel is not restricted to paired days.
daily_necb_long <- final_daily %>%
  select(Date, year, LD_NEE_gC_day, MD_NEE_gC_day) %>%
  pivot_longer(
    cols = c(LD_NEE_gC_day, MD_NEE_gC_day),
    names_to = "Orchard",
    values_to = "NEE_gC_m2_day"
  ) %>%
  mutate(
    Orchard = recode(
      Orchard,
      "LD_NEE_gC_day" = "LD",
      "MD_NEE_gC_day" = "MD"
    ),
    Orchard = factor(Orchard, levels = c("LD", "MD"))
  ) %>%
  filter(is.finite(NEE_gC_m2_day))

harvest_events <- harvest_C %>%
  transmute(
    Date = HarvestDate,
    Orchard = factor(Orchard, levels = c("LD", "MD")),
    NECB_increment_gC_m2 = Harvest_gC_m2,
    Increment_type = "Harvested-fruit export"
  )

daily_events <- daily_necb_long %>%
  transmute(
    Date,
    Orchard,
    NECB_increment_gC_m2 = NEE_gC_m2_day,
    Increment_type = "Daily NEE"
  )

# Add a zero starting point so both trajectories visibly begin at zero.
start_rows <- tibble(
  Date = as.Date("2023-12-31"),
  Orchard = factor(c("LD", "MD"), levels = c("LD", "MD")),
  NECB_increment_gC_m2 = 0,
  Increment_type = "Starting value"
)

figure3_necb_support <- bind_rows(
  start_rows,
  daily_events,
  harvest_events
) %>%
  group_by(Orchard, Date) %>%
  summarise(
    NECB_increment_gC_m2 = sum(NECB_increment_gC_m2, na.rm = TRUE),
    Increment_type = paste(sort(unique(Increment_type)), collapse = "; "),
    .groups = "drop"
  ) %>%
  arrange(Orchard, Date) %>%
  group_by(Orchard) %>%
  mutate(
    Cumulative_NECB_gC_m2 = cumsum(NECB_increment_gC_m2)
  ) %>%
  ungroup()

# Weekly difference uses paired daily dates because it directly compares orchards.
# Difference = MD - LD:
#   positive -> LD had greater uptake or lower CO2 release;
#   negative -> MD had greater uptake or lower CO2 release.
figure3_weekly_support <- final_daily %>%
  filter(
    is.finite(LD_NEE_gC_day),
    is.finite(MD_NEE_gC_day)
  ) %>%
  mutate(
    Daily_difference_MD_minus_LD = MD_NEE_gC_day - LD_NEE_gC_day,
    Week = floor_date(Date, unit = "week", week_start = 1)
  ) %>%
  group_by(Week) %>%
  summarise(
    Paired_days = n(),
    Weekly_difference_MD_minus_LD = sum(
      Daily_difference_MD_minus_LD,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  mutate(
    Advantage = case_when(
      Weekly_difference_MD_minus_LD > 0 ~ "LD",
      Weekly_difference_MD_minus_LD < 0 ~ "MD",
      TRUE ~ "Equal"
    ),
    Bar_color = case_when(
      Advantage == "LD" ~ unname(orchard_colors["LD"]),
      Advantage == "MD" ~ unname(orchard_colors["MD"]),
      TRUE ~ "grey60"
    )
  )

write_csv(
  figure3_necb_support,
  file.path(figure_support_dir, "Figure_3_chained_NECB_support.csv")
)

write_csv(
  figure3_weekly_support,
  file.path(figure_support_dir, "Figure_3_weekly_NEE_difference_support.csv")
)

figure3_endpoints <- figure3_necb_support %>%
  group_by(Orchard) %>%
  slice_max(Date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(Orchard, Date, Cumulative_NECB_gC_m2)

print(figure3_endpoints, n = Inf)

# Growing-season shading and labels.
growing_seasons <- gs_dates %>%
  transmute(
    year,
    xmin = gs_start,
    xmax = gs_end,
    xmid = gs_start + as.integer((gs_end - gs_start) / 2),
    label = paste0(year, "\nGrowing season")
  )

harvest_labels <- harvest_C %>%
  distinct(HarvestDate) %>%
  transmute(
    Date = HarvestDate,
    label = "Harvest"
  )

x_limits <- c(as.Date("2023-12-15"), as.Date("2025-12-31"))
month_breaks <- seq(
  as.Date("2024-01-01"),
  as.Date("2025-12-01"),
  by = "1 month"
)

necb_range <- range(
  figure3_necb_support$Cumulative_NECB_gC_m2,
  na.rm = TRUE
)

necb_span <- diff(necb_range)
if (!is.finite(necb_span) || necb_span == 0) necb_span <- 1

upper_label_y <- necb_range[2] - 0.06 * necb_span
harvest_label_y <- necb_range[2] - 0.18 * necb_span

weekly_range <- range(
  figure3_weekly_support$Weekly_difference_MD_minus_LD,
  na.rm = TRUE
)

weekly_positive_y <- max(0, weekly_range[2]) * 0.72
weekly_negative_y <- min(0, weekly_range[1]) * 0.72

if (!is.finite(weekly_positive_y) || weekly_positive_y == 0) {
  weekly_positive_y <- 1
}

if (!is.finite(weekly_negative_y) || weekly_negative_y == 0) {
  weekly_negative_y <- -1
}

p_figure3_top <- ggplot(
  figure3_necb_support,
  aes(
    x = Date,
    y = Cumulative_NECB_gC_m2,
    color = Orchard
  )
) +
  geom_rect(
    data = growing_seasons,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.55
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.5
  ) +
  geom_vline(
    data = harvest_labels,
    aes(xintercept = Date),
    inherit.aes = FALSE,
    linetype = "dotted",
    linewidth = 0.55
  ) +
  geom_line(linewidth = 1.05, na.rm = TRUE) +
  geom_text(
    data = growing_seasons,
    aes(x = xmid, y = upper_label_y, label = label),
    inherit.aes = FALSE,
    family = "Helvetica",
    fontface = "bold",
    size = 3.8,
    lineheight = 0.9,
    color = "black"
  ) +
  geom_text(
    data = harvest_labels,
    aes(x = Date, y = harvest_label_y, label = label),
    inherit.aes = FALSE,
    family = "Helvetica",
    angle = 90,
    vjust = -0.5,
    size = 3.2,
    color = "black"
  ) +
  scale_color_manual(
    values = orchard_colors,
    breaks = c("LD", "MD")
  ) +
  scale_x_date(
    limits = x_limits,
    breaks = month_breaks,
    labels = function(x) lubridate::month(x),
    expand = c(0, 0)
  ) +
  labs(
    tag = "a",
    x = NULL,
    y = "Chained cumulative NECB (g C m⁻²)",
    color = "Orchard"
  ) +
  publication_theme(base_size = 11) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = c(0.94, 0.13),
    legend.justification = c(1, 0),
    legend.background = element_rect(
      fill = "white",
      color = "grey70"
    ),
    panel.grid.major.y = element_line(color = "grey88", linewidth = 0.35)
  )

p_figure3_bottom <- ggplot(
  figure3_weekly_support,
  aes(
    x = Week,
    y = Weekly_difference_MD_minus_LD
  )
) +
  geom_rect(
    data = growing_seasons,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "grey85",
    alpha = 0.55
  ) +
  geom_col(
    aes(fill = Bar_color),
    width = 6,
    color = "black",
    linewidth = 0.12
  ) +
  scale_fill_identity() +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.5
  ) +
  geom_vline(
    data = harvest_labels,
    aes(xintercept = Date),
    inherit.aes = FALSE,
    linetype = "dotted",
    linewidth = 0.55
  ) +
  annotate(
    "text",
    x = as.Date("2024-02-01"),
    y = weekly_positive_y,
    label = "LD greater uptake\nor lower CO₂ release",
    hjust = 0,
    family = "Helvetica",
    fontface = "bold",
    size = 3.1,
    lineheight = 0.9,
    color = "black"
  ) +
  annotate(
    "text",
    x = as.Date("2024-02-01"),
    y = weekly_negative_y,
    label = "MD greater uptake\nor lower CO₂ release",
    hjust = 0,
    family = "Helvetica",
    fontface = "bold",
    size = 3.1,
    lineheight = 0.9,
    color = "black"
  ) +
  scale_x_date(
    limits = x_limits,
    breaks = month_breaks,
    labels = function(x) lubridate::month(x),
    expand = c(0, 0)
  ) +
  labs(
    tag = "b",
    x = "Month of year",
    y = "Weekly NEE difference, MD − LD\n(g C m⁻² wk⁻¹)"
  ) +
  publication_theme(base_size = 11) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(color = "grey88", linewidth = 0.35)
  )

p_figure3 <- p_figure3_top / p_figure3_bottom +
  plot_layout(heights = c(3.5, 1.35))

print(p_figure3)

save_publication_figure(
  p_figure3,
  basename = "Figure_3_Chained_Cumulative_NECB_Weekly_NEE_Difference_2024_2025",
  width = 7.2,
  height = 7.1
)

```

# Figure captions and provenance

```{r captions_provenance}
# ============================================================
# 7) CAPTIONS, PROVENANCE, AND SESSION INFORMATION
# ============================================================

rain_total_2024 <- figure2_totals %>%
  filter(as.character(Year) == "2024") %>%
  pull(Total_precipitation_mm) %>%
  round(0)

rain_total_2025 <- figure2_totals %>%
  filter(as.character(Year) == "2025") %>%
  pull(Total_precipitation_mm) %>%
  round(0)

figure2_caption <- paste0(
  "Figure 2. Cumulative growing-season precipitation measured at the ",
  "low-density (LD) orchard eddy-covariance tower in 2024 and 2025. ",
  "Curves begin at budbreak on 28 March 2024 and 15 April 2025 and extend ",
  "through 31 October. Total growing-season precipitation was ",
  rain_total_2024, " mm in 2024 and ", rain_total_2025, " mm in 2025. ",
  "Although precipitation accumulated more slowly during much of June through ",
  "mid-August 2025, several late-season rainfall events resulted in the greater ",
  "2025 total."
)

figure3_caption <- paste0(
  "Figure 3. Chained cumulative net ecosystem carbon balance (NECB; upper ",
  "panel) and weekly differences in net ecosystem exchange (NEE; lower panel) ",
  "for the low-density (LD) and moderate-density (MD) apple orchards during ",
  "2024 and 2025. Cumulative NECB was calculated by continuously accumulating ",
  "daily NEE across both years and adding harvested-fruit carbon as an ecosystem ",
  "export at each harvest, indicated by the vertical dotted lines. The 2025 ",
  "trajectories therefore begin from the ending 2024 carbon balances rather ",
  "than resetting to zero. Negative cumulative NECB values indicate net carbon ",
  "retention within the orchard ecosystem. Gray shading denotes the growing ",
  "seasons. Lower-panel bars represent weekly MD − LD NEE differences. Positive ",
  "values indicate greater carbon uptake or lower CO₂ release by LD, whereas ",
  "negative values indicate greater carbon uptake or lower CO₂ release by MD. ",
  "Dashed horizontal lines indicate zero."
)

caption_lines <- c(
  figure2_caption,
  "",
  figure3_caption
)

writeLines(
  caption_lines,
  con = file.path(figure_main_dir, "Figure_2_and_Figure_3_captions.txt")
)

figure_provenance <- tribble(
  ~Figure, ~Primary_source, ~Calculation,
  "Figure 2",
  basename(figure2_workbook),
  "Growing-season cumulative sum of P_RAIN_tower_mm using documented phenology dates",
  "Figure 3 upper panel",
  paste(basename(unname(daily_paths)), collapse = "; "),
  "Orchard-specific daily U50 NEE accumulated continuously across 2024-2025; harvested-fruit carbon added as export",
  "Figure 3 lower panel",
  paste(basename(unname(daily_paths)), collapse = "; "),
  "Weekly sum of paired daily MD minus LD NEE differences"
)

write_csv(
  figure_provenance,
  file.path(figure_support_dir, "figure_provenance.csv")
)

capture.output(
  sessionInfo(),
  file = file.path(figure_main_dir, "sessionInfo_manuscript_figures.txt")
)

cat(
  "\nManuscript figure workflow completed successfully.",
  "\nFigure files written to: ", figure_main_dir,
  "\nSupport files written to: ", figure_support_dir,
  "\n",
  sep = ""
)
```
